### 2017-08-24 paper.tex
Update on Overleaf.

\documentclass[useAMS,usenatbib]{mnras}

\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cite}
\usepackage{natbib}
\usepackage{xcolor}
\usepackage{import}

\def\aap{Astronomy \& Astrophysics}
\def\apj{The Astrophysical Journal}
\def\mnras{Monthly Notices of the Royal Astronomical Society}
\def\aj{Astronomical Journal}
\def\aaps{AAPS}
\def\nat{Nature}

\usepackage{newtxtext,newtxmath}
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{hyperref}

\newcommand{\fl}[1]{{\color{magenta}FL: #1}}
\newcommand{\red}[1]{{\color{red} #1}}
\newcommand{\degree}{^{\circ}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{etoolbox}
\makeatletter
\newcount\c@additionalboxlevel
\setcounter{additionalboxlevel}{0}
\newcount\c@maxboxlevel
\setcounter{maxboxlevel}{1}
\patchcmd\@combinedblfloats{\box\@outputbox}{%
  \stepcounter{additionalboxlevel}%
  \box\@outputbox
}{}{\errmessage{\noexpand\@combinedblfloats could not be patched}}

\AtBeginShipout{%
  \ifnum\value{additionalboxlevel}>\value{maxboxlevel}%
    \typeout{Warning: maxboxlevel might be too small, increase to %
      \the\value{additionalboxlevel}%
    }%
  \fi 
  \@whilenum\value{additionalboxlevel}<\value{maxboxlevel}\do{%
    \typeout{* Additional boxing of page `\thepage'}%
    \setbox\AtBeginShipoutBox=\hbox{\copy\AtBeginShipoutBox}%
    \stepcounter{additionalboxlevel}%
  }%
  \setcounter{additionalboxlevel}{0}%
}
\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

\title[SL Challenge]{The Strong Gravitational  Lens Finding Challenge}
%C.~E.~Petrillo$^{1}$,  C.~Tortora$^{1}$, L.~V.~E.~Koopmans$^{1}$, \and G.~Verdoes~Kleijn$^{1}$
\author[R. B. Metcalf {\it et al.}]{
R. Benton Metcalf,$^{1,4}$\thanks{E-mail: robertbenton.metcalf@unibo.it} 
M. Meneghetti,$^4$ 
Fabio Bellagamba,$^1$
Emmanuel Bertin,$^3$ 
\newauthor
R\'emi Cabanac,$^2$ 
Etienne Decenci\`ere,$^{10}$
R\'emi Flamary,$^8$
Raphael Gavazzi,$^{17}$ 
\newauthor
Mario Geiger,$^6$ 
Philippa Hartley,$^7$
Marc Huertas-Company,$^9$
Neal Jackson,$^7$ 
\newauthor
Eric Jullo,$^{16}$
Jean-Paul Kneib,$^6$
Leon Koopmans,$^{22}$
Carlo Enrico Petrillo,$^{22}$ 
Christoph Sch\"{a}fer,$^6$ 
Alessandro Sonnenfeld,$^5$ 
\newauthor
Amit Tagore,$^7$
Crescenzo Tortora,$^{22}$
Diego Tuccillo,$^{9,10}$
Santiago Velasco-Forero,$^{10}$
Gijs Verdoes Kleijn,$^{22}$ 
Camille Avestruz,$^{11,12,13,\dagger}$
\newauthor
Nan Li,$^{14}$
Matthew Lightman,$^{15}$
Cl\'ecio R. Bom,$^{18,19}$
Manuel B. Valent\'in,$^{19}$
Martin Makler,$^{19}$
\newauthor
Fran\c{c}ois Lanusse,$^{20}$ Quanbin Ma,$^{21}$ Chun-Liang Li,$^{21}$
\newauthor
and ?\\
% List of institutions
$^{1}$ Departimento di Fisica \& Astronomia, Universit\`a di Bologna, via Gabetti 93/2, 4012 Bologna, Italy \\
$^{2}$ IRAP, Universit\'e de Toulouse, CNRS, UPS, Toulouse, France.\\
$^{3}$ Sorbonne Universit\'{e}s, UPMC Univ. Paris 6 and CNRS, UMR 7095, Institut d'Astrophysique de Paris,  98bis Bd Arago, F-75014,\\ Paris France.\\
$^{4}$ INAF-Osservatorio Astronomico di Bologna, via Ranzani 1, 40127 Bologna, Italy \\
$^{5}$ Kavli IPMU (WPI), UTIAS, The University of Tokyo, Kashiwa, Chiba 277-8583, Japan\\
$^{6}$ \red{EPFL}\\
$^{7}$ Jodrell Bank Centre for Astrophysics, School of Physics \& Astronomy, University of Manchester, Oxford Rd, Manchester M13 9PL, UK \\
$^{8}$ Laboratoire Lagrange, Universi\'{e} de Nice Sophia-Antipolis, Centre National de la Recherche Scientifique, Observatoire de la C\^{o}te d'Azur, \\ Parc Valrose, 06108 Nice, France \\
$^{9}$ LERMA, Observatoire de Paris, CNRS, Universit\'e Paris Diderot, 61, Avenue de l'Observatoire F-75014, Paris, France \\
$^{10}$ MINES Paristech, PSL Research University, Centre for Mathematical Morphology, 35 rue Saint-Honore, Fontainebleau, France \\
$^{11}${Enrico Fermi Institute, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^{12}${Kavli Institute for Cosmological Physics, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^{13}${Department of Astronomy \& Astrophysics, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^{14}${School of Physics and Astronomy, Nottingham University, University Park, Nottingham, NG7 2RD, UK
}\\
$^{15}${JPMorgan Chase, Chicago, IL 60603 U.S.A.}\\
$^{16}$\red{LAM} \\
$^{17}$\red{IAP} \\
$^{18}${Centro Federal de Educa\c{c}\~ao Tecnol\'ogica Celso Suckow da Fonseca, CEP 23810-000,  Itagua\'i, RJ, Brazil}\\
$^{19}${Centro Brasileiro de Pesquisas F\'isicas, CEP 22290-180, Rio de Janeiro, RJ, Brazil}\\
$^{20}${McWilliams Center for Cosmology, Department of Physics, Carnegie Mellon University, Pittsburgh, PA 15213, USA}\\
$^{21}${School of Computer Science, Carnegie Mellon University, Pittsburgh, PA 15213, USA}\\
$^{22}${Kapteyn Astronomical Institute, University of Groningen, Postbus 800, 9700 AV, Groningen, The Netherlands}\\
$^\dagger${Provost's Postdoctoral Scholar at the University of Chicago}\\
}

\begin{document}

\date{Accepted . Received ; in original form }

\maketitle

\label{firstpage}

% Abstract of the paper
\begin{abstract}
We present a description and results of an open gravitational lens finding challenge.  Participants were asked to classify 100,000 candidate objects as to whether they were gravitational lenses or not with the goal of developing better automated methods for finding lenses in large data sets.   
\end{abstract}

\begin{keywords}
gravitational lensing -- cosmology 
\end{keywords}

\section{Introduction}
\label{sec:introduction}

Strong gravitational lenses are rare cases where a distant galaxy or quasar is aligned so closely with a foreground galaxy or cluster of galaxies  that the gravitational field of the foreground object creates multiple, highly distorted images of the background object.  These take the form of multiple images of a quasar and Einstein rings surrounding a galaxy and of highly distorted arcs seen through galaxy groups and clusters.  The first strong lens was discovered in 1979 \citep{1979Natur.279..381W} and since then many valuable scientific uses have been found for them.  They have been used to study how dark matter is distributed in galaxies and clusters \citep[e.g.][]{1991ApJ...373..354K,2001ApJ...554.1216C,2002ApJ...568L...5K,2003ApJ...587..143R,2003ApJ...583..606K,2005MNRAS.360.1333W,2005ApJ...623...31D,2009MNRAS.392..945V,2016MNRAS.463.3115T}.  They have been used to measure the Hubble constant and other cosmological parameters \citep[e.g.][]{1964MNRAS.128..307R,1992ARA&A..30..311B,2000ApJ...544...98W,2013ApJ...766...70S,2016A&ARv..24...11T}.
Their magnification has been used as a natural telescope to observe otherwise undetectable objects \citep[e.g][]{2007ApJ...671.1196M,2017MNRAS.464.4823B,2016ApJ...833..264S}.   They have put limits on the self interaction of dark matter and on the theory of gravitation.  Through microlensing they have been used to study the structure of quasars \citep{2008ApJ...689..755M,2008ApJ...673...34P,2011ApJ...729...34B}.   To access this wealth of information the first step is to find the lenses.  \red{Could use some citations in this paragraph.}

The Square Kilometer Array (SKA)\footnote{http://skatelescope.org/}, the Large Synoptic Survey Telescope (LSST)\footnote{https://www.lsst.org/} and the Euclid space telescope\footnote{https://www.euclid-ec.org/} are expected to increase the number of potential lenses by orders of magnitude \citep{collett_15,euclidSLWGwhitepaper,2015aska.confE..84M,2010MNRAS.405.2579O}.   For example it is estimated that there will be approximately 200,000 observable galaxy-galaxy lenses in the Euclid data set among \red{tens of billions (better number??)} of potential objects .  So far, less than a thousand lenses have been found in total across many heterogeneous data sets.  These surveys will bring a new era for strong lensing where large relatively well defined samples of lenses will be possible.  It will also require handling much larger quantities of data than has been customary in this field.

Up to this point, the most widely used method for finding lenses in imaging surveys has been by visual inspection of a samples candidates that have been selected on the bases of luminosity and/or color.  This has been done in the radio \citep{2003MNRAS.341...13B} and in the visible with space and ground based data \citep{2008MNRAS.389.1311J,2008ApJS..176...19F,2014MNRAS.439.3392P} \red{(Maybe there are some others to cite here?)}.  A related method pioneered by the SPACE WARPS project \citep{2016MNRAS.455.1171M,2016MNRAS.455.1191M,2015MNRAS.452..502G} has been to crowd source the visual inspection through an online platform.  These efforts have proved very fruitful, but they deal with orders of magnitude fewer candidates and lenses than will be necessary in the future.  Dealing with such large quantities of data will not be practical for any visual inspection approach.  In addition, the efficiency and detection bias of human inspection methods are difficult to rigorously quantify.

Galaxy scale lenses have also been searched for spectroscopically by looking for high redshift stellar lines in the spectra of lower redshift large galaxies.  Notably this was done in the Sloan Lens ACS (SLACS) survey, producing a relatively well defined and pure sample of Einstein ring lenses \citep{2006ApJ...638..703B}.  However, spectroscopy is telescope-time consuming and for the foreseeable future we are not likely to have spectroscopic surveys that cover anywhere near the number of objects as the planned imaging surveys.

Some automated algorithms have been developed in the past to detect lenses in images.  These have been designed to detect arc-like features \citep{2006astro.ph..6757A,2007A&A...472..341S} and rings \citep{2014ApJ...785..144G,2014A&A...566A..63J}.  They have been applied to survey data and found of order 200 lenses \citep{2007A&A...461..813C,2012ApJ...749...38M,2016A&A...592A..75P}.  Another approach has been to use a lens modeling code to fit a model to all the candidates and pick out the ones that fit the model well \citep{2009ApJ...694..924M,2010A&A...517A..25S,2017arXiv170401585S}.

More recently machine learning techniques that have become widely used in the fields computer image processing and artificial intelligence 
have been applied to this problem and others in astronomy, in particular, artificial neural networks (ANNs) and support vector machines (SVM). These algorithms do not pick out particular predefined features, but learn from a training set which features are most important for distinguishing categories of objects.  These have been used widely for such tasks as handwriting and facial recognition.  In astronomy they are beginning to be used for categorizing galaxy morphologies \citep{2015MNRAS.450.1441D}, photometric redshifts \citep{2017MNRAS.465.1959C,2016PASP..128j4502S,2017NewA...51..169S}, supernova classification \citep{2016ApJS..225...31L} and the lens finding problem \citep{2017MNRAS.471..167J,2017arXiv170207675P,2017MNRAS.465.4325O,2017A&A...597A.135B}.

Given the future of this field with large amounts of data coming soon and many new ideas emerging, it is timely to stage a series of challenges to stimulate new work, determine what can realistically be done in lens finding and get a better idea of the strengths and weaknesses to different methods.  The long term goal is to get a set of algorithms that can handle Euclid, LSST or SKA data sets and produce high purity and high completeness lens samples with well defined efficiency or selection.  We anticipate further challenges in the future in which the realism of the data simulations will become progressively better.  Here we have chosen to concentrate on galaxy/small group scale lenses where the background source is a galaxy because we feel that this is where the most progress can be made and the scientific return is the highest, but QSO lenses and cluster/group lenses challenges may follow.

The paper is organized as follows.  The form of the challenge and its rules are described in the next section.  The methods used to simulate mock images of galaxies and gravitational lenses are described in section~\ref{sec:simulation}.  In section~\ref{sec:entries} each of the methods that were used to solve the challenge are briefly described.  We discuss the metrics used to evaluate entries in section~\ref{sec:figure_of_merit}.  The performance of each of the methods broken down in different ways is presented in section~\ref{sec:performance}.  Finally,  in section~\ref{sec:conclusion}, we conclude with a discussion of what was learned and how methods can be improved in the future.

\section{The Challenge}
\label{sec:challenge}

The challenge was in fact two separate challenges that could be entered independently.  One was designed to mimic a single band of a future imaging data set from a satellite survey such as Euclid \citep{***} \red{(the latest description)}.  The other was designed to mimic ground based data with multiple bands, roughly modeled on the Kilo-Degree Survey (KiDS)\footnote{http://kids.strw.leidenuniv.nl/} \citep{2013ExA....35...25D}. In neither case were the simulated images meant to precisely mock these surveys, but the surveys were used as guides to set noise levels, pixel sizes, sensitivities, and other parameters.

In each case, a training set of 20,000 images in each band was provided for download at any time along with a key giving some properties of the object including whether it was a gravitational lens.  Each image was $101\times101$ pixels.  The researchers were free to download these sets and train their algorithms on them.  To enter the contest the participants needed to register with a team name at which point they would be given a unique key and the address of a test data set.  These data sets contained 100,000 candidates.  In the case of the multi-band ground-based set this was 400,000 images. The participant had 48 hours to upload a classification of all candidates consisting of a score between 0 and 1 -- 0 signifying the lowest confidence that it is a lens and 1 signifying the highest confidence that it is a lens.  This ranking could have been a simple binary (0 or 1) classification  or it could have been a continuous range of numbers representing the probability of being a lens or it could have been a finite number of confidence levels.

The challenge was opened on November 25, 2016 and closed on February 5, 2017.

\section{the simulations}
\label{sec:simulation}

Creating the mock images starts with a cosmological N-body simulation.  In this case it was the Millennium simulation \citep{2009MNRAS.398.1150B}.   A catalog of dark matter halos within a light-cone is constructed within the Millennium Observatory project \citep{2013MNRAS.428..778O}.   The challenge sets were based on a 1.6 sq.deg. light cone extending out to redshift $z=6$.
The halo catalog is read into the GLAMER lensing code \citep{2014MNRAS.445.1942M,2014MNRAS.445.1954P} to do all the ray-tracing.  This code projects the halos onto lens planes, in this case 20, and finds the deflection angle on each plane by a hybrid tree method.  In this way the halos have the mass, concentration and clustering properties from the N-body simulation, but within each strong lens the mass resolution is not limited by the original simulation, but follows the analytic mass profile. {\color{red} MM: I am a bit confused: at this stage, how are the halos modeled? N-body or analytic? What is the shape of the halos?}

With GLAMER we identify and map out all the caustics within the light-cone for 33 source planes -- z=1 to 3 in intervals of 0.1 and 3 to 6 in intervals of 0.25.  We take every caustic that corresponded to a critical curve with an Einstein radius larger than 1.5 times the resolution of the final images.  The Einstein radius is estimated here and in all that follows as $R_{\rm ein}=\sqrt{A_{\rm ein}/\pi}$ where $A_{\rm ein}$ is the angular area within the critical curve.

 A **** square degree light-cone gives rise to **** such caustics, but these over count the number of lenses since most will have caustics on multiple source planes.  These lenses could be used as is, but we wanted to produce a much larger number with more randomness.   For each caustic we identify the lens plane with the highest convergence and identify all the halos within a three dimensional distance of 0.5 Mpc from the center of the critical curve and on this and its neighboring lens planes.    This collection of halos is then used as the lens and rotated to produce more random lenses.    It  contains all the sub-halos and nearby companion halos, but not the large scale structure surrounding it.

For lensed sources we use sources from the Hubble Ultra Deep Field (UDF) that have been decomposed into shapelet functions to remove noise.  This is the same set of images as used 
in  \citep{2008A&A...482..403M,2010A&A...514A..93M,2016arXiv160604548M}.   There are 9,350 such sources with redshifts and separate shapelet coefficients in 4 bands.

To construct a mock lens, first a caustic on the highest redshift source plane is selected.  This is done in order of Einstein area, but all the critical curves are used more than once.  Since every lens with a caustic at a lower redshift will have a caustic at the highest redshift this is a selection from all of the caustics in the light-cone.   The lens is extracted as explained above and rotated randomly in three dimensions.  A source is selected at random from the shapelet catalog subject to a magnitude limit in a reference band.   The redshift of the UDF source is used as the source redshift.  If the source is at a lower redshift than the lens or within $\Delta z = 0.4$ another random source is selected.

The furthest point in the caustic is found from its own center and the source is placed randomly within 3 times this distance.  This is a somewhat arbitrary length designed to be a compromise between producing only very clear strong lenses because all the sources are right in the center of the caustic and making the process  inefficient because most of the sources are too far away from the caustic to produce clear lenses.  If the source positions where taken completely at random the fraction of clear lenses would be very low.

The visible galaxies associated with the lens must also be simulated. There  too few bright galaxies in the  in the UDF catalog to make enough mock lens galaxies for this purpose.  Instead, for most of the lenses, we used an analytic model for the surface brightness of these galaxies.  The Millennium Observatory provides parameters for the galaxies that inhabit the dark matter halos using the semi-analytic galaxy formation models of \citet{2011MNRAS.413..101G}.  The parameters used here were the total magnitude, the bulge-to-disk ratio, the disk scale height and the bulge effective radius.  The magnitude and bulge-to-disk ratio are a function of the pass band.  Each galaxy is given a random orientation and inclination angle.between 0 and $80\degree$.  The disk is exponential with no vertical height which is why the inclination is limited to $80\degree$.  The bulge is represented by a S\'{e}rsic profile with a axis ratio randomly sampled between 0.5 and 1.  The S\'{e}rsic index, $n_s$, is given by 
\begin{align}
\log( n_s )= 0.4 \log\left[ {\rm max}\left(\frac{B}{T},0.03\right)\right] + 0.1 x 
\end{align}
\red{(reference)} where $\frac{B}{T}$ is the bulge to total flux ratio and $x$ is a uniform random number between -1 and 1.  This approximately reproduces the observed correlation between these quantities.

In addition to the basic disk and bulge models we introduce some spiral arms.  The surface brightness of 
the disks are given by
\begin{align}
S(\theta,r) = e^{-r/R_h} \left( 1 + A \cos(N_a\theta + \phi_r \right) ~, \\
\phi_r = \alpha\log(2 r/ R_h) + \phi_d \nonumber
\end{align}
where $R_h$ is the scale height of the disk.  The phase angle of the arms, $\phi_d$, is chosen at random.  The parameters $A$, $\alpha$ and $N_a$ are chosen from distributions that are judged by eye to produce realistic 
galaxies.
The bulge is also perturbed from a perfect S\'{e}rsic profile by multiplying the surface brightness by
\begin{align}
1+\sum_{n=1}^6 a_n \cos\left( n \theta  + \phi_n \right)
\end{align}
where $\phi_n$ is a random phase.  The coefficients are picked randomly from between -0.002 and 0.002. \red{(reference)}

These foreground galaxies are rotated in three dimensions with the halos of the lens each time a random lens is produced so that they remain in the same positions relative to the mass.  All the random parameters are also reassigned with every realization of the lens.

These images of the of foreground galaxies are not intended to reproduce the true population of galaxies, but only to be sufficiently irregular to make them difficult to fit to a simple analytic model that might make them very easy to distinguish from a foreground plus a lensed image.   As will be discussed later, more realistic models will be needed in the future and are a subject of current investigation.

To represent the mass of the galaxies we make a gridded map of the surface brightness at 3 times the resolution of the final image.  The surface brightness map in converted into a mass map by assuming a uniform mass-to-light ratio of 1.5 solar in the reference band.  This mass map is added to the lens before the ray tracing is done. {\color{red} MM: Again, I am confused with how the lenses are modeled in this stage of the simulation and during the search for caustics described in the beginning of this section.}

The code is able to produce any combination of foreground galaxies, lensed image and noise that is desired.  For the training set we provided an image of the total lens with noise, an image of the foreground galaxies with noise and image of the lensed background source without noise.
For the test sets only the final images were provided to participants although all the information was stored for analyzing the entries.

\subsection{Space-based}
\label{sec:sim-space-based}

The space-based datasets were meant to roughly mimic the data quality which is expected from observations by the Euclid telescope in the visible channel. To this end, the pixel size was set to 0.1 arcsec and a Gaussian PSF was applied with a FWHM of 0.18 arcsec. The Gaussian PSF is clearly a simplified model, but a realistic treatment of the PSF is outside the scope of this paper. The typical width of the blurring due to the PSF is anyway realistic. The reference band for background and foreground galaxies was SDSS $i$, which is overlapping with the broader Euclid VIS band. The realisation of the mock images followed the same procedure described in \citet{2004PASP..116..750G} and \citet{2008A&A...482..403M}. As a result, the noise follows a Gaussian distribution with a realistic width and is uncorrelated between pixels. Characteristics of the instrument, filter and exposure times were taken from the Euclid Red Book \citep{2011arXiv1110.3193L}.

 In the challenge set the limiting magnitude for background sources to 28 in $i$.   60\% of the cases had no background source and were thus labeled as non-lenses.

\subsection{Ground-based}
\label{sec:sim-ground-based}

For the ground-based images four bands ($u,g,r$, and $i$) where simulated {\color{red} (these are SDSS filters too, right?)}.  The reference band was $r$.  For the challenge set 85\% images where made with purely simulated images as outlines above and the other 15\% used actual images taken from a preliminary sample of bright galaxies directly from the KiDS survey.  Lensed source images where added to these real images at the same rate as for the mock images,  in this case 50\%.  No attempt was made to match the halo masses to the observed galaxies in these cases.  These real images where added for more realism and so that, by comparing the results for real and mock images, we can evaluate how realistic our simulations are in this context.  There were about 160,000 of these stamps from KiDS.

The KiDS survey provided a representative PSF map in each band that was applied to all mock images.   The pixel size in this case was 0.2~arcsec. 
Weight maps for the KiDS images were also provided.  Some of these had masked regions from removed stars, cosmic rays, and bad pixels.  For the mock images the noise was simulated by adding normally distributed numbers with the variance given by the weight maps.  The weight maps were also randomly rotated and flipped.  This resulted in many of the images having large masked regions in them.

By chance one of the original KiDS images appears to have been a lens.  When an additional lensed source was added this made a double lens or "jackpot" lens.

\section{The Entries}
\label{sec:entries}

There were 24 valid entries into the challenge.  They are listed in table~\ref{table:entries}.  There were a variety of different methods used and participants come from a variety of different backgrounds, most were professional astronomers, but there were also entries from people outside of the field.

\begin{table*}
\centering
\begin{tabular}{rlll}
  \hline
 & Name & type & authors \\ 
  \hline
  1 & AstrOmatic & Space-Based & Bertin   \\ 
  2 & GAHEC IRAP & Space-Based & Cabanac  \\ 
  3 & CAS Swinburne Melb & Ground-Based & Jacobs   \\ 
  4 & ALL-star & Ground-Based & Avestruz, N.Li \& Lightman  \\ 
  5 & Manchester1 & Space-Based & Jackson, Hartley \& Tagore  \\ 
  6 & CMU-DeepLens-Resnet-Voting & Space-Based & Ma, Lanusse \& C. Li  \\ 
  7 & Manchester SVM & Ground-Based & Hartley   \\ 
  8 & CMU-DeepLens-ResNet & Space-Based & Francois Lanusse, Ma, C. Li \& Ravanbakhsh \\ 
  9 & CMU-DeepLens-Resnet-Voting & Ground-Based & Ma, Lanusse \& C. Li   \\ 
  10 & YattaLensLite & Space-Based &  Sonnenfeld  \\ 
  11 & NeuralNet2 & Space-Based &  Davies  \\ 
  12 & CAST & Ground-Based &  Roque De Bom, Valent\'{\i}n \&  Makler  \\ 
  13 & CMU-DeepLens-ResNet-ground3 & Ground-Based &  Lanusse, Ma, Ravanbakhsh \& C. Li \\ 
  14 & GAMOCLASS & Space-Based & Marc Huertas-Company, Tuccillo, Velasco-Forero \& Decenci\~{a}re \\ 
  15 & LASTRO EPFL (CNN) & Space-Based & Geiger, Sch\"{a}fer \& Kneib \\ 
  16 & Manchester SVM & Space-Based &  Hartley  \\ 
  17 & CMU-DeepLens-ResNet-aug & Space-Based & Ma,  Lanusse, Ravanbakhsh \& C. Li \\ 
  18 & LASTRO EPFL & Ground-Based & Geiger, Sch\"{a}fer \& Kneib   \\ 
  19 & CAST & Space-Based & Roque De Bom, Valent\'{\i}n \& Makler  \\ 
  20 & AstrOmatic & Ground-Based & Bertin  \\ 
  21 & All-now & Space-Based & Avestruz, N. Li \& Lightman   \\ 
  22 & Manchester-NA2 & Ground-Based & Jackson \& Tagore  \\ 
  23 & YattaLensLite & Ground-Based & Sonnenfeld   \\ 
  24 & Kapteyn Resnet & Space-Based & Petrillo, Koopmans, Kleijn, Tortora \& Vernardos \\ 
   \hline
\end{tabular}
\caption{Entries to the challenges.}
\label{table:entries}
\end{table*}

\section{lens finding methods}
\label{sec:methods}

This section contains short descriptions of the lens finding methods that were used in 
the challenge.  Each subsection refers to a team which gave a separate entry.

\subsection{GAHEC IRAP (Cabanac)}

\begin{figure}
 \includegraphics[width=\columnwidth]{figures/arcmethod.pdf}
 \caption{ (GAHEC IRAP) From top-left to bottom right, 1) a simulated arc extracted from SL challenge in which an tuned Arcfinder selects 3 candidates (green circles), 2) the smoothed image on which pixel wise elongation is computed, 3) the resulting elongated pixels after threshold, 4) the set of pixels selected for the computation of arc candidate properties. }
 \label{fig:Cabanac}
\end{figure}

Arcfinder \citep{2006astro.ph..6757A,2007A&A...461..813C,2012ApJ...749...38M} illustrated Figure \ref{fig:Cabanac}, is a fast linear method that computes a pixel wise elongation parameter (ratio of first-order moments in a n-pix window oriented in proper reference frame) for all pixels of mexican-hat-smoothed FITS images. Arcfinder then extracts contiguous pixels above a given background and computes the candidate arcs length, width, area, radius of curvature and peak surface brightness. A final thresholding is set to maximize purity over completeness on a few typical arcs of the dataset.
For the current SL challenge, arcfinder was tuned to detect long and narrow arcs, and was optimized on a subset of 1000 simulated images with a grid covering a range of elongation windows and arc areas.  A python wrapper allows users to change parameters in a flexible way and run the arcfinder C code as a linux line command. Arcfinder took a couple of hours to run on the entire dataset with some overheads due to the dataset format. The code is publicly available at https://github.com/rcabanac/arcfinder.

\subsection{AstrOmatic (Bertin)}

The lens detector is based on a convolutional neural network (CNN), trained with the provided training datasets. The CNN is implemented in Python, using the TensorFlow framework\footnote{http://www.tensorflow.org/}. Both ground multichannel and space monochannel image classifiers have the exact same CNN architecture.

{\color{red} (MM: I think that the co-authors should provide a more detailed description of the algorithms. For example, they should quickly explain what is a CNN and what are its general components (i.e. what kind of layers, activation functions, etc.)}

The network itself consists of three convolutional layers ($11\times 11\times 32$, $5\times 5 \times 64$ and $3\times 3\times 64$), followed by two fully-connected layers ($256$ and $64$ neurons) and an output softmax layer {\color{red} MM: the numbers make little sense without a description of what the mean -- see my previous comment}. The first five layers use the ELU (Exponential Linear Unit) activation function \citep{2015arXiv151107289C}, which in our tests led to significantly faster convergence compared to ReLU and even SoftPlus activation. Dropout regularization \citep{2012arXiv1207.0580H,JMLR:v15:srivastava14a} is applied to both convolutional and fully connected layers, with ``keep'' probabilities $p=2/3$ and $p=1/2$, respectively.

Prior to entering the first convolutional layer, input image data are rescaled and dynamic-range compressed with function $f(x) =
\mathrm{arcsinh} (10^{11} x)$, and bad pixels are simply set to 0.
Data augmentation is performed in the form of random up-down and left-right image flipping, plus $k\pi/2$ rotations, where $k$ is a random integer in the $[0,3]$ range. Additionally, a small rotation with random angle $\theta$ is applied, involving bicubic image resampling. $\theta$ follows a Gaussian distribution with mean $\mu=0$ and standard deviation $\sigma_{\theta}=5^{\circ}$. No attempt was made to generate and randomize bad pixel masks in the data augmentation process.

The CNN weights are initialized to random values using a truncated Gaussian distribution with mean $\mu=0$ and standard deviation $\sigma=5.10^{-2}$. The network is trained on a Titan-X ``Pascal'' nVidia GPU using the \textit{Adam} gradient-based optimizer \citep{2014arXiv1412.6980K} during 800 epochs, with an initial learning rate $\eta(t=0)=10^{-3}$ and a learning rate decay $\eta(t+1)/\eta(t)=0.99$, where $t$ is the epoch. Because of a lack of time, tests were limited to assessing the basic classification performance on a subset of the of 1,000 images/datacubes, using the 19,000 others for training.

{\red MM: May be we could separate the methods into three categories: machine-learning, arc-finders, visual inspection and group them accordingly}

\subsection{YattaLens Lite (Sonnenfeld)}
YattaLensLite is a simpler version of the algorithm YattaLens \citep{2017arXiv170401585S}, modified to meet the time constraints of the challenge.
YattaLensLite subtracts a model surface brightness profile describing the lens galaxy from the $g$-band image, then runs SExtractor to detect tangentially elongated or ring-shaped objects, which are interpreted as lensed images.
In the ground-based challenge, the model lens surface brightness profile is obtained by taking a rescaled version of the $i$-band image.
The difference in color between lens and source usually allows the lensed images to still be detectable after the lens subtraction process.
However, in order to avoid subtracting off the lensed images in systems with similar colors between lens source, we radially truncate the model lens surface brightness.
The model lens light is truncated at the smallest radius between the position where the surface brightness is comparable to the sky background level, or the position of a positive radial gradient in surface brightness, if detected.

In the space-based challenge, it is not possible to separate lens and source based on color, because only data in one band is provided. The lens light model then is produced by taking a centrally-inverted image and then using the same truncation prescription used with ground-based data. The central inversion step is taken to reduce the chances of subtracting flux from lensed images, which are in general not centrally symmetric as opposed to typical lens galaxies.

In the full version of YattaLens, a lens modeling step is performed to improve the purity of the sample. However, such a procedure is too time consuming and was not performed in this challenge.

\subsection{LASTRO EPFL (Geiger, Sch\"{a}fer)}

We used a convolutional neural network \citep{Fukushima1980,Lecun1998} with a simple architecture of 12 layers (inspired by \citep{symmetry}), see table \ref{tab:architecture}.
To avoid the problem of the data flow distribution getting out of the comfort zone of the activation functions (Internal Covariate Shift), we used a mix of normalization propagation \citep{norm_prop} (without the constraint on the weights but a proper initialization) and batch normalization \citep{batch_norm} (slowly disabled over the iterations).
As activation function, we used a scaled and shifted ReLU (Rectifier Linear Unit), 
\begin{equation} \label{eq:relu}
    \frac{1}{\sqrt{\pi-1}} (\sqrt{2 \pi} \max(0,x) - 1),
\end{equation}
to satisfy the properties required by the normalization propagation.
In our implementation of batch normalization, instead of using only the moments computed over the current batch, we used a low-pass filter ($\bar\mu_i \longleftarrow (1-\eta) \; \bar\mu_{i-1} + \eta \; \mu_i(\text{batch})$, $\eta$ set to $1$ at the beginning and decay with the iterations).
For the training, the 20k provided images were split into two sets, 17k for training and 3k for validation.
Each iteration of the gradient descent (more precisely \textit{Adam} \citep{adam}) minimizes the cross entropy, 
\begin{equation} \label{eq:xent}
    \left\{
    \begin{array}{ll}
        - \log(p)   & \text{if the image is a true lens} \\
        - \log(1-p) & \text{if the image is a nonlens}
    \end{array}
    \right.,
\end{equation}
where $p$ is the output of the neural network, computed over a batch of 30 images, 15 lenses and 15 nonlenses, picked from the training set.
Making batches of 30 images adds noise to the gradient makes the iterations faster.

To augment the training set, each image of the batch is transformed with a random transformation of the dihedral group (rotations of 90 degrees and mirrors), its pixel values multiplied by a factor picked between $0.8$ and $1.2$ and shifted by a random value between $-0.1$ and $0.1$.
To prevent the overfitting, we used some dropout \citep{dropout} (with a keeping probability decreasing with the iterations).
The masked region of the ground based images are handled by simply setting them to zero.
Each final prediction is made of the product of the predictions of the 8 transformations of the image by the dihedral group.
The architecture is implemented in Tensorflow\footnote{\url{http://tensorflow.org/}}.
Our code is accessible on github\footnote{\url{https://github.com/antigol/lensfinder-euclid}}.

\begin{table*}
    \centering
    \begin{tabular}{|l|l|l|l|}
        \hline
        Layer type & shape & activation & \# parameters \\ \hline \hline

        \textbf{convolutional 4x4} & $101\!\times\! 101\!\times\!1/4 \to 98\!\times\!98\!\times\!16$ & rectifier & 256/1'024 + 16 \\ \hline
        \textbf{convolutional 3x3} & $98\times98\times16 \to 96\times96\times16$ & rectifier & 2'304 + 16 \\ \hline
        max pool /2 & $96\times96\times16 \to 48\times48\times16$ & - & - \\ \hline
        batch normalization & $48\times48\times16$ & - & 16 + 16 \\ \hline

        \textbf{convolutional 3x3} & $48\times48\times16 \to 46\times46\times32$ & rectifier & 4'608 + 32 \\ \hline
        \textbf{convolutional 3x3} & $46\times46\times32 \to 44\times44\times32$ & rectifier & 9'216 + 32 \\ \hline
        max pool /2 & $44\times44\times32 \to 22\times22\times32$ & - & - \\ \hline
        batch normalization & $22\times22\times32$ & - & 32 + 32 \\ \hline

        \textbf{convolutional 3x3} & $22\times22\times32 \to 20\times20\times64$ & rectifier & 18'432 + 64 \\ \hline
        \textbf{convolutional 3x3} & $20\times20\times64 \to 18\times18\times64$ & rectifier & 36'864 + 64 \\ \hline
        max pool /2 & $18\times18\times64 \to 9\times9\times64$ & - & - \\ \hline
        batch normalization & $9\times9\times64$ & - & 64 + 64 \\ \hline
        dropout & $9\times9\times64$ & - & - \\ \hline

        \textbf{convolutional 3x3} & $9\times9\times64 \to 7\times7\times128$ & rectifier & 73'728 + 128 \\ \hline
        dropout & $7\times7\times128$ & - & - \\ \hline
        \textbf{convolutional 3x3} & $7\times7\times128 \to 5\times5\times128$ & rectifier & 147'456 + 128 \\ \hline
        batch normalization & $5\times5\times128$ & - & 128 + 128 \\ \hline
        dropout & $5\times5\times128$ & - & - \\ \hline

        \textbf{fully-connected} & $5\times5\times128 \to 1024$ & rectifier & 3'276'800 + 1'024 \\ \hline
        dropout & $1024$ & - & - \\ \hline
        \textbf{fully-connected} & $1024 \to 1024$ & rectifier & 1'048'576 + 1'024 \\ \hline
        dropout & $1024$ & - & - \\ \hline
        \textbf{fully-connected} & $1024 \to 1024$ & rectifier & 1'048'576 + 1'024 \\ \hline
        batch normalization & $1024$ & - & 1'024 + 1'024 \\ \hline
        \textbf{fully-connected} & $1024 \to 1$ & sigmoid & 1'024 + 1 \\ \hline \hline
        Total & - & - & $\approx$ 5'674'000 \\ \hline
    \end{tabular}
    \caption{LASTRO EPFL architecture}
    \label{tab:architecture}
\end{table*}

\subsection{Manchester/Manchester1 (Jackson,Tagore)}

All images (a total of 100000) were examined for each of the space- and
ground-based datasets. This was done by two observers; AT examined 30000
images in each case and NJ examined 70000. Observation was carried out
over a 48-hour period, at the rate of 5000/hr (NJ) and 2500/hr (AT). 
The overall results, in terms of ROC curves, were very similar for 
both observers. The space-based challenge produced areas of 0.800 and 
0.812 for NJ and AT respectively, and the ground-based challenge yielded 
0.891 and 0.884.

The Python scripts used for manual examination of multiple images are available
onGitHub\footnote{https://github.com/nealjackson/bigeye} and are described in more detail
in \citet{hartley2017support}. For one-colour data such as
the space-based training set, the images are individually colour-scaled using
square-root scaling. The bright limit of the colour-scale is determined
from the pixel values in a rectangle comprising the inner ninth of the
image area, with the limit being chosen as the $n$th centile of the pixel
values in this area. Values between $n=95$ and $n=98$ give optimum results,
judging by experiments on the training set. The number of images in each
grid was also optimised using the training set, with 16$\times$8 or 
8$\times$4 giving good results on one-colour data. For three-colour data,
such as the ground-based challenge data, the individual bands for each 
object are colour-scaled and then combined into an RGB image. In this case
8$\times$4 grids were used for examination, due to the generally lower 
resolution of the images. The script also allows the user to adjust the
colour-scale in real time when examining and marking images, and records
the image name corresponding to the image within which the cursor resides
at the time any key is pressed, together with the key.

Images were classified by both observers into 5 categories, ranging from
0 (no evidence of any lensed objects in the image) to 4 (certain lenses).
For both observers, the rate of false positives in the ``certain'' lenses
was between 0.1\% and 0.3\%. The exception was the ground-based imaging 
for one observer, where a 4.6\% rate resulted mainly from a
single decision to allow a false-positive ``double lens'' which occurred
repeatedly throughout the data at different orientations. The false-negative
rate among the class-0 identifications was similar for both observers, at
around 25\% for the space-based images and 20\% for the ground-based.

\subsection{Gabor-SVM (Hartley, Flamary)}

A Support Vector Machine (SVM) is a supervised machine learning method which uses labeled training data to determine a classification model (see e.g., \citet{vapnik79estimation}, \citet{Cortes1995} and \citet{Burges1998}). A preprocessing stage first extracts a set of useful features from input samples, before projecting each sample as a vector in multidimensional space. The model then separates classes of data by maximising the margin between a defining hyperplane and a set of so-called support-vectors at the inner edge of each class. The process of classification is computationally inexpensive since optimisation depends only on the dot products of the support vector subset. Feature extraction, however,  requires both an extensive exploration of the feature space during the development of a model, and potentially intensive computer resources in order to transform the original samples. Our full method is described in detail in \citet{hartley2017support}, and was developed using the Python scikit-learn and scikit-image packages \citep{scikit-learn,scikit-image}.

During our development of an SVM classifier for lens finding, feature extraction initially involved the decomposition of each image into a set of objects, using SExtractor \citep{1996A&AS..117..393B} and GALFIT \citep{2002AJ....124..266P} to recover and subtract objects iteratively. This method had previously been used in a static algorithm approach which assigned points according to the morphological properties of each image \citep[see][]{2014A&A...566A..63J}. Lensed-like objects displaying, for example, greater ellipticity and tangentiality were awarded more points. Since the SVM operates in a fixed dimensional space,  properties of individual objects were collapsed into a fixed set describing the mean and variance of morphological properties of all the objects within an image. After training an SVM using these features we recorded a modest separation of lens and non-lens classes.

An alternative approach was to design a set of Gabor filters to be applied to each sample. The Gabor kernel is described by a sinusoidal function multiplied by a Gaussian envelope. We discard the imaginary part of the function to leave, in two-dimensional space:

\begin{equation}
G_c[i,j]=Be^{-\frac{(i^2+j^2)}{2\sigma^2}} \mathrm{cos}\left(\frac{2\pi}{\lambda} (i\, \mathrm{cos} \, \theta + j\, \mathrm{sin} \,\theta)\right),
\end{equation}

where harmonic wavelength $\lambda$, Gaussian spread  $\sigma$ and orientation $\theta$ define the operation performed on each point $i,j$ in an image. Such a kernel is a popular image processing choice for edge detection and texture classification \citep[e.g.][]{Feichtinger98a,Springer-verlag97computationalmodels} and is thought to mimic some image processing functions of the mammalian brain \citep{Jones1233}.

\begin{figure}
  \centering
      \includegraphics[width=1\columnwidth]{figures/polarfilter.pdf} 
  \caption{Example of our feature extraction procedure used to transform a ring. The image on the right shows the response of a set of Gabor filters after convolution with a polar transformed image of an Einstein ring. The strongest response is seen in the orientation perpendicular to the radial direction and at the frequency most closely matching that of the ring.}
 \label{gaborring}
\end{figure}

Our final feature extraction procedure first applied a polar transform to each image in order to exploit the edge detection of the Gabor filter, picking out tangential components typical of galaxy-galaxy lensing. Each image was then convolved with several Gabor filters of varying frequency and rotation (see Fig.~\ref{gaborring}). Stability selection methods were used to investigate the classification performance using different combinations of filters. The responses for each Gabor filter when applied to each image were measured by calculating statistical moments for each filtered image. These moments formed  our final input data on which the SVM could be trained and applied. We used brute-force optimisation methods to select a non-linear SVM containing radial basis function (RBF) kernel and tuned a small set of regularisation hyperparameters to achieve good generalisation performance. During training and testing, our final and best scores achieved when testing on the lens finding challenge training data were an AUC of 0.88 for the space set and 0.95 for the ground set. Classification was performed using a modest desktop PC.

\subsection{GAMOCLASS (Tuccillo, Huertas-Company, Velasco-Forero, Decenci\`ere)}

\begin{figure}
  \centering
      \includegraphics[width=1\columnwidth]{figures/Fig1_gamoclass.pdf} 
  \caption{GAMOCLASS schematic}
 \label{Fig1_gamoclass}
\end{figure}

Our algorithm for classification of strong lensed galaxies was developed as a convolutional neural network (CNN) used for binary classification task.
Our code was trained and tested on the space-based dataset (dataset 0) provided for the Strong lensing challenge, comprising 20,000 stamps of single band simulated galaxies. We used the full dataset in the proportion of 4/5 for training and 1/5 for validation. The training images were labelled with 1 if showing strong lensing and 0 otherwise. Our CNN gives as output a probability [0,1] of the input image being a strongly lensed system.
The final architecture of our model is illustrated in Fig. \ref{Fig1_gamoclass}. The input image (101x101 pixels) is first processed by a 2D convolution layer with a 3x3 filter size, then subsampled by a 3x3 max pooling layer. Other two identical units follow, with a growing dimensionality of the output space (depth, i.e.output number of filters) in the convolution, for a total of 3 convolutional layers and 3 max pooling layers. Each of these convolutional layers is followed by a rectified linear unit (ReLU) step. The output of these units is then processed through a single fully-connected layer follower by a dropout layer, and, finally, by a one-neuron fully connected layer with sigmoid activation function. For the classification problem we used the binary cross-entropy cost function and for weights updates we used an adaptive moment estimation (ADAM, \citealt{Kingma_2014}) optimization method. The use of the ADAM optimizer improves the learning rate, and previous tests with stochastic gradient descent (SGD) lead to worse results on the classification of the validation sample.
In order to increase the size of the training set and make the model invariant to specific transformations, we perform several data augmentation steps:
\begin{itemize}
\item Random rotation of the image in the range [0, 180$\degree$], using a reflection fill mode to keep constant the size of the images.
\item The images were randomly shifted of 0.02 times the total width of the image.
\item Random horizontal and vertical flips of the images
\end{itemize}
During the training we initialize the weights of our model with random normal values and we warm up the training \citet{Huang_2016} of the CNN for 25 epochs, using an exponential decay rate (1e-6) and a staring learning rate of α = 0.001. Then the network is trained using an early stopping method, and for a maximum number of 300 epochs. The early stopping method is an effective method to prevent overfitting and consists in stopping the training if a monitored quantity does not improve for a fixed number (called \textit{patience}) of training epochs. The quantity that we monitored was the \textit{accuracy} of the classification of the validation sample. The best architecture was trained over 220 epochs with a parameter of patience equal to 20.
We implemented our code the Keras framework \citet{Chollet_2015} on top of Theano \citet{Bastien_2012}.
Our architecture converges with a classification accuracy of 91\% on the validation sample. We further evaluated the performance of our classifier calculating the ROC curve of the classifier, i.e. the True Positive Rate (TPR) against the False Positive Rate. As shown in Fig. \ref{Fig2_gamoclass}, we reach a TPR higher than the 90\% with a FPR < 8\%.

\begin{figure}
  \centering
      \includegraphics[width=1\columnwidth]{figures/Fig2_gamoclass.pdf} 
  \caption{GAMOCLASS ROC on training data.}
 \label{Fig2_gamoclass}
\end{figure}

%-------------------------------------------------%
\subsection{ALL Methodology (Avestruz, Li, Lightman}
%-------------------------------------------------%

We detail the ALL team methodology in \citet{avestruz_etal17}. The
pipeline was originally developed to automatically classify strong
lenses in mock Hubble Space Telescope and Large Synoptic Sky Telescope
data generated with code described in \citet{li_etal16} and
\citet{collett_15}.  We apply the same exact steps for the single-band
data for Euclid, but modify the feature extraction step for the
four-band KIDS data.  We summarize the steps below.

We use tools from {\em Scikit-learn} \citep{pedregosa_etal12} and
perform minimal image preprocessing. First, we replace masked pixels
with the average of surrounding pixels, then enhance contrast in the
image by taking the normalized log of pixel values.  The next step of
our pipeline consists of a feature extraction stage, where our feature
vector is a {\em histogram of oriented gradients} (HOG)
\citep{dalalandtriggs_05} that quantifies edges in the image.  HOG has
three main parameters that determine the binning and resolution of
edges captured by the features.  The result is a one dimensional
feature vector corresponding to the magnitude of oriented gradients
across the image.  With the KIDS data, we extract a feature vector for
each of the four bands and concatenate the vectors to create a final
feature vector for each object that we use to train a model
classifier.

We use {\em Logistic Regression} (LR), a machine learning algorithm,
to train a classifier model.  LR requires a parameter search over the
regression coefficient, $C_{LogReg}$.  The parameters from both the
feature extrator, HOG, and the linear classifier, LR, contain
parameters that we optimize for peak model performance.  We use {\em
  GridSearchCV} from {\em Scikit-learn} to select cross-validated
parameters for HOG parameters and a subset of $C_{LogReg}$ values with
20\% of the test images provided.  We then run a finer parameter
search over $C_{LogReg}$, splitting the test images into 80\% training
and and 20\% test to avoid overfitting the data.  We use the best
parameters to then train the entire dataset for the final model
classifier that we used to evaluate the competition data.

%-------------------------------------------------%
\subsection{CAST (Bom, Valent\'in, Makler)}
%-------------------------------------------------%
[Note: we are placing this section here, but we agree it is good to group the methods. Perhaps machine learning based methods could go on a separate section and a unified description can be given to CNN-based methods, as most follow under this category.]

The CBPF Arc Search Team (CAST) tested several arcfinding schemes using as core a CNN. For both the space-based and ground-based samples we carried out a simple preprocessing phase to enhance the objects in the images in order to check if they improve the automated arc detection with CNN. We chose a contrast adjustment with $0.1\%$ pixel saturation and apply a low pass band Wiener filter \citep{wiener1964extrapolation} to reduce the effect of the noise.

We used a native CNN from Matlab\footnote{\texttt{https://www.mathworks.com/products/matlab.html}, \texttt{https://www.mathworks.com/help/nnet/convolutional-neural-networks.html}}, which has Convolutional 2D layers with 20 5x5 filters. This CNN can work either with one or three input images, representing greyscale and colour images.
Therefore, we employed different strategies for the two samples available for the challenge, which involve combinations of the available bands running in one or more CNN, using or not the preprocessing, and combining the output with the aid of other machine learning methods.
In each case we used the simulations made available for the challenge both to train and to validate the results and we used the area under the ROC to determine which combination of methods gives the best result.    
We selected 
%We evaluate the ROC curves for all configurations tested, selecting 
$90\%$ of the images, chosen randomly, for the training and $10\%$ to validate. We repeated the process $10$ times to avoid bias due a specific choice of training/validation set and to define an uncertainty in our ROC.

For the space-based data set we tested only two configurations: i) using the CNN straightforwardly for classification and ii) with the preprocessing described above. 
We found that the results, accounting for the uncertainties,  
were clearly superior for the area under the ROC in the case 
including the preprocessing. Therefore, this is the configuration we have used for the challenge.

As mentioned above, the CNN used can take 3 images as input. Thus, if we want to use the information on the $4$ available bands, we will have to either combine $2$ of the $4$ bands to end up with $3$ bands for a single RGB CNN (configuration I below) or we use multiple CNNs to use the information available (configuration II to VI below). To combine the outputs of several CNNs we use a Support Vector Machine \citep[hereafter SVM; see e.g., ][]{rebentrost2014quantum} also implemented in Matlab. The SVM is used to combine the outputs $p_i$ of the several CNN (configurations II, III, IV and VI). Instead of using only $p_i$ as inputs to the SVM we also tested providing the SVM with image features obtained by the CNN (feature maps, configuration V) as inputs. In all cases we tested with and without the preprocessing

A more detailed description of each configuration tested is presented below:
\begin{enumerate}
\item Combination of bands r and i with the average between bands u and g. Use one CNN for classification.
\item Creation of $1$ CNN for each band (total of $4$). The $4$ outputs are used as input to a SVM classifier which returns the final classification $p$. 
\item Combination into 4 different combinations of bands:  RGB $\rightarrow$ (\textit{u,g,r}), (\textit{u,g,i}), (\textit{u,r,i}) and (\textit{g,r,i}). One CNN for each combination of bands and then use of the output score as input to an SVM classifier.
%Mediatrix features used as input of an SVM classifier.
%
%IV) Use of wavelet decomposition for each band then use of these decomposed images in the same combinations of case III.
\item  Average of bands in different combinations RGB $\rightarrow$ (\textit{ug,r,i}), (\textit{u,gr,i}) and (\textit{u,g,ri}). The outputs of these 3 CNNs are inputs to a SVM classifier.
\item Use of CNN-activations (CNN feature maps) as inputs to a SVM classifier, using same combinations of bands of III. The output of each CNN is used as input to a SVM classifier.
\item Use of wiener filter and contrast adjustment on each band, then using the resulting images in the same architecture as in (iii). 
\end{enumerate}

For the ground based cases three configurations with highest area under ROC were III,IV and VI. Although the areas are very similar between IV and VI the last one is superior in the low fake positives end. Thus, for the Strong Lensing Challenge in the ground base sample, we used configuration VI. This final scheme is ilustraded in figure \ref{colorCASTCNN}.

The area under ROC, in both space based configurations were, in general, smaller than in the multi-band case, which suggests how the CNNs are sensitive to color information to find Strong Lensing. Particularly,  ground base configuration II used one CNN per band thus not considering color information and has the similar area under ROC as our best single band configuration.

%In total we tested XX configurations for the ground based challenge.

%Based on the ROC, we picked for the blind challenge XX, which is illustrated in figure \ref{colorCASTCNN}.

\begin{figure}
 \includegraphics[width=\columnwidth]{figures/SLchallengeCAST.png}
 \caption{CAST Lens finder pipeline for the ground based sample. Illustration of the chosen architecture for the CAST search in the case of ground-based simulated images.}
 \label{colorCASTCNN}
\end{figure}

\subsection{CMU DeepLens (Lanusse, Ma, Li, Ravanbakhsh)}

\subsubsection{Convolutional Neural Networks}
\fl{I'm placing this section here but it's intended as a global introduction to CNNs. Feel free to move/modify/discard as you want. Also I tried to keep things concise but some concepts may still require a little bit explanations, let me know.}

A Convolutional Neural Network (CNN) \citep{Lecun1998} is a multi-layer feed-forward neural network model particularly well suited for processing natural images. With the very recent advances of the \textit{Deep Learning} framework \citep{Lecun2015}, models based on CNN architectures have reached or even surpassed human accuracy in image classification tasks \citep{He2015a}.

The fundamental building block of a CNN is the \textit{convolutional layer}. This element applies a set of convolution filters on an input image to produce a series of so-called \textit{feature maps}. The coefficients of these filters are free parameters that are learned by the model. We will use the notation 
$n \times n - n_c$ to denote a convolutional layer using filters of size $n \times n$ pixels and outputting $n_c$ feature maps \fl{As Massimo pointed out, I think it's useful to introduce this notation and harmonize its use}. In typical architectures, the size of these convolution filters is kept small (i.e. 3x3 or 5x5 pixels) to limit the complexity of the model. To capture features on larger scales, CNNs rely on a multi-resolution approach by interleaving convolutional layers with \textit{pooling layers} which apply a downsampling operation to the feature maps. The most common downsampling schemes are the max-pooling and average pooling strategies, which downsample an input image by taking respectively the maximum or average values within a given region (e.g. 2x2 patches for a downsampling of factor 2).

A CNN architecture is therefore a stack of convolution layers and pooling layers, converting the input image into an increasing number of feature maps of progressively coarser resolution. The final feature maps can capture information on large scales and can reach a high-level of abstraction. To perform the classification itself from these feature maps, the CNN is typically topped by a fully-connected neural network outputting the class probability of the input image.

For a binary classification problem such as the one involved in strong lens detection, the training is performed by optimizing the weights of the model as to minimize the \textit{binary cross-entropy} objective:
\begin{equation}
	- \sum_{n=1}^{N} y[n] \log \hat{y}[n] + (1 - y[n])\log(1 - \hat{y}[n]) \;,
\end{equation}
where $N$ is the number of training instances, $y \in \{0,1\}$ is the true class of the image and $\hat{y} \in [0,1]$ is the class probability predicted by the model. This optimization is performed by a Stochastic Gradient Descent (SGD) algorithm which updates the model iteratively by taking small gradient steps over randomly selected subsamples of the training set (so called \textit{mini-batches}). All the CNN-based methods presented in this work rely on the ADAM \citep{Kingma_2014} algorithm, an accelerated variant of SGD algorithm.

\fl{Things I still have to mention: activation function, dropout, batch normalisation}.

% Francois: This is an attempt at comparing the different CNN-based methods, not sure it's really useful
%\begin{table*}
%\begin{tabular}{|c|c|c|c|c|c|c|}
%\hline 
%Method       & Depth & Activation & Optimizer & Regularisation &   Pre-processing &  learning rate\\ 
%\hline 
%AstrOmatic &  3 + 2 & ELU             & ADAM       &   Dropout            &   asinh                   &     0.001  +decay\\
%LastroEPFL &   12    & ReLU           &  ADAM      &  Batch norm       & rotations and stuf  & 0.001 + decay\\
%GLAMOCLASS & 3 + 1 & ReLU    &  ADAM      &   dropout last layer & rotations and stuff & 0.001 + decay \\
%CMU DeepLens	& 47  & ELU 	    	 &  ADAM       &  batch norm     & rotations and stuff & 0.001 + decay drop \\
%\hline 
%\end{tabular} 
%\caption{Comparison of CNN based methods}
%\end{table*}

\subsubsection{CMU DeepLens}

\texttt{CMU DeepLens} is based on a residual network (or \textit{resnet}) architecture \citep{He2015a}, a modern variant of CNNs which can reach much greater depths (over 1000 layers) while still gaining accuracy. We provide a short overview of our model below but a full description of our architecture can be found in \citet{Lanusse2017}.

Much like conventional CNNs, resnets are based on convolutional and pooling layers. However, resnet differ from CNNs by the introduction of so-called shortcut connections bypassing blocks of several convolutional layers. As a result, instead of learning the full mapping from their input to their output these residual blocks only have to learn the difference to the identity. In practice, this difference allows residual networks to be trained even for very deep models. For a more thorough description of this architecture, we refer the interested reader to Section~2.3 of \citet{Lanusse2017}.

Our baseline model is composed of a first 7x7-32 convolutional layer which can accommodate either single-band or multi-band images. The rest of the model is composed of 5 successive blocks, each block being made of 3 resnet units (specifically, pre-activated bottleneck residual units \citep{He2016}). At each block, the signal is downsampled by a factor 2 and the number of feature maps is in turn multiplied by 2. The model is topped by an average-pooling layer followed by a single fully-connected sigmoid layer with a single output. Apart from the final layer, we use the ELU activation throughout. The weights of the model are initialized using random normal values, following the strategy advocated in \citet{He2015a}.

Training was performed using the ADAM optimizer with mini-batches of size 128 over 120 epochs, with an initial learning rate of $\alpha=0.001$, subsequently decreased by a factor 10 every 40 epochs. This multi-step training procedure is important to progressively refine the model parameters of the model and achieve our final accuracy.

We adopt a minimal pre-processing pre-processing strategy for the input images, removing the mean image and normalising by the noise standard deviation $\sigma$ in each band, these statistic being evaluated over the whole training set. In addition, we clip extreme values above $250 \sigma$ to limit the dynamic range of the input. Bad-pixels are simply set to 0 after this pre-processing step.

Given the relatively small training set preventing overfitting is an important consideration. In our final model, we combine several data augmentation strategies: random rotations (in the range $[-90, \ 90^\circ]$), random mirroring along both axes, and random resizing (by a small factor in the range $[0.9, \ 1]$).

The architecture presented above is the one that lead to our best results in both branches of the challenge, i.e. \texttt{CMU-DeepLens-Resnet} for space-based and \texttt{CMU-DeepLens-Resnet-ground3} for ground-based. We also submitted results for two variants of this baseline model, named \texttt{-aug} and \texttt{-Voting}. 
The first variant introduced several data-augmentation schemes, including the ones mentioned above and the addition of Gaussian noise to the input images. We found however that the introduction of noise was not necessary as the other methods were enough to prevent overfitting. 
The second variant was used to explore a voting strategy between three different models. These models differed by the type of residual blocks (bottleneck vs wide) and by their handling of missing pixels (setting to 0 or to the median value of the image). The predictions of the best 2 out of 3 models were then averaged to produce the final classification probability.

Our model is implemented using the \texttt{Theano}\footnote{\url{http://deeplearning.net/software/theano/}} and \texttt{Lasagne}\footnote{\url{https://github.com/Lasagne/Lasagne}} libraries. On an Nvidia Titan Xp GPU, our full training procedure requires approximately 6 hours on the ground-based challenge, but classification of the whole testing set is performed in a couple of minutes. Finally, in the interest of reproducible research, our code is made publicly available at
\begin{center}
\url{https://github.com/McWilliamsCenter} 
\end{center}
This repository also contains a notebook detailing how to reproduce our challenge submission.

\subsection{Kapteyn Resnet (Petrillo, Tortora, Verdoes Kleijn, Koopmans) }
Our lens-finder is based on a Convolutional Neural Network (CNN), following the strategy adopted recently in \citep{2017arXiv170207675P}.

We decide to treat the problem as a three-class classification problem where the classes are \textit{non-lenses}, \textit{clear lenses} and \textit{dubious lenses}. We define the \textit{dubious lenses} as the lenses with lensing features with less than 160 pixels and the \textit{clear lenses} those with more than 160 pixels belonging to the lensed source. This choice is motivated by the fact that specializing the network in recognizing different class could lead to a more robust classification. In addition, in a hypothetical application of the method to real data from a survey, this could be a way to select the most blatant lenses.

Given the rules of the challenge, we had to submit a prediction for each source in the test-set as a single number: a number in the interval [0,1] or either a 0 or a 1. Unfortunately, to satisfy this requirement we had to collapse one of the class into another: we provided a 0 when the source is classified as a \textit{non-lens} and 1 if it is classified as a \textit{clear lenses} or as a \textit{dubious lenses}. Moreover, this did not allow to build a continuous ROC curve but only a binary one. \textbf{(We need to test if this was an improvement or not. I have the submission ready.)}

Ihe final submission is made using the average values of the predictions from the same network trained three different times.

\subsubsection{Implementation}
The CNN is implemented in Python 2.7 using the open-source libraries \textsc{Lasagne}\footnote{\href{http://github.com/Lasagne/Lasagne/}{\tt http://github.com/Lasagne/Lasagne/}} 
and \textsc{Theano}
\footnote{\href{http://deeplearning.net/software/theano/}{\tt http://deeplearning.net/software/theano/}} \citep{theano}.  
The training of the CNN is executed on a GeForce GTX 760 in parallel with the data augmentation performed on the CPU using the \textsc{scikit-image}\footnote{\href{http://scikit-image.org/}{\tt http://scikit-image.org/}} package \citep{van2014scikit}.

\subsubsection{Architecture}
We used the CNN architecture called ResNet described in \citep{he2015deep} with three stacks of residual blocks of 5 layers.

The output layer is composed by three units. Each units gives in output a number between 0 and 1 that represents, respectively, the probability of being a non-lens, a small-lens, a big-lens.

\subsubsection{Preprocessing}
The training fits files are preprocessed with the software \textsc{STIFF}\footnote{\href{http://www.astromatic.net/software/stiff}{\tt http://www.astromatic.net/software/stiff}} which automatically converts the fits files to gray-scale TIFF images operating a non-linear intensity transformations to enhance the low-brightness features of the image.

Due to memory limitations we down-sampled the images to 84 by 84 pixels. Most likely, this leads to a worse classification accuracy than using the original image resolution.

We augmented the training images in the following way: i) random rotation of 90, 180 or 270 degrees; ii) random shift in both $x$ and $y$ direction between -2 and +2 pixels; iii) $50\%$ probability of horizontally flipping the image.

Finally, the image border is cropped in order to have 80 by 80 pixel input images.

\subsubsection{Training}
The network is trained by minimizing the categorical cross-entropy loss function
\begin{equation}
L = -\sum_j t_{j}\log p_{j}
\label{EQloss}
\end{equation}
where the $t_j$ and $p_j$ are respectively the label and the prediction for the class $j$. 
The minimization is done via mini-batch stochastic gradient descent with \textit{Adam} updates \citep{Kingma_2014}. We used a batch size of 56 and performed  46000 gradient updates. We started with a learning rate of $4 \times 10^{-4}$, decrease it to $4 \times 10^{-5}$ after 35000 updates and to $4 \times 10^{-6}$ after 43000 updates.
The initial values of the biases are set to zero and the weights of each filter are initialized from a random normal distribution with variance ${2/n}$ where $n$ is the number of inputs of the unit (see \citep{he2015delving}).  
We use L2-norm regularization with $\lambda= 9 \times 10^{-3}$.

\section{results}
\label{sec:results}

In this section we summarize some of our analysis of the submissions.  In section~\ref{sec:figure_of_merit} we discuss how to judge a classifier in this particular case and define some metrics of success.

\subsection{figures of merit}
\label{sec:figure_of_merit}

In deriving a good figure of merit for evaluating lens finding algorithms one needs to take into account the particular nature of this problem.
The traditional method for evaluating a classification algorithm is with the receiver operating characteristic curve, or {\bf ROC} curve.  This is a plot of the true positive rate (TPR) and the false positive rate (FPR).  In this case these are defined as
\begin{align}
{\rm TPR} &= \frac{\textrm{ number of true lenses classified as lenses}}{\textrm{ total number of true lenses}} \\
{\rm FPR} &= \frac{\textrm{number of non-lenses classified as lenses}}{\textrm{ total number of non-lenses}}
\end{align}
The classifier generally gives a probability of a case being a lens, $p$ in which case a threshold is set and everything with $p$ greater is classified as a lens and everything smaller is classified as not a lens.  The TPR and FPR are then plotted as a curve parametrised by this threshold.  At $p=1$ all of the cases are classified as non-lenses and so TPR=FPR=1 and at $p=0$ all of the cases are classified as lenses so TPR=FPR=0.  These points are always added to the ROC curve.  If the classifier made random guesses then the ratio of lenses to non-lenses would be the same as the ratio of the number of cases classified as lens to the number of cases classified as non-lenses and so TPR=FPR.  The better a classifier is the smaller the FPR and the larger the TPR so the further away from this diagonal line it will be.  When a classifier provides only a binary classification or a discrete ranking, the ROC connects the endpoints to the discrete points found be using each rank as a threshold.

A common figure of merit for a classifier is the area under the ROC ({\bf AUROC}).  This evaluates the overall ability of a classifier to distinguish between cases.  This was the criterion on which the challenge participants were told to optimize.  However, in the case of gravitational lensing this is not the only thing, and not the most important thing, to consider.  Gravitational lenses are rare events, but to improve the discrimination and training of the classifiers the fraction lenses in test and training sets are boosted to something around half.  In these circumstances it is important to consider the absolute number of cases that will be missclassified when the fraction of true cases is closer to what is expected in the data.

If the rates of false positives and false negatives remain the same in real data  the contamination of the sample will be
\begin{align}
\frac{\rm FP}{\rm TP} \simeq \frac{\rm FPR}{\rm TPR} \left(\frac{\textrm{number of non-lenses in sample}}{\textrm{number of lenses is sample}} \right)
\end{align}
Since only about one in a thousand objects will be lenses (perhaps somewhat more depending on pre-selection) the contamination will be high unless the FPR is much less than the TPR.  For this reason we consider some additional figures of merit.

The {\bf TPR$_0$} will be defined as the highest TPR reached, as a function of $p$ threshold, before a single false positive occurs in the test set of 100,000 cases.  This is the point were the ROC meets the FPR = 0 axis.  This quantity highly penalizes classifiers with discrete ranking which often get TPR$_0$ = 0 because their highest classification level is not conservative enough to eliminate all false positives.  We also define {\bf TPR$_{10}$} which is the TPR at the point were less than ten false positives are made.  If the TP rate is boosted from the FPR by a factor of 1,000 in a realistic data set this would correspond to about a 10\% contamination.

In addition to these considerations, the performance of a classifier is a function of many characteristics of the lens system.  It might be that one classifier is good at finding systems with large Einstein radii and incomplete arcs, but not as good at finding small complete Einstein rings that are blended with the light of the lens galaxy.  Also lens may have a source that is too faint to be detected by any algorithm or is too far from the lens to be very distorted, but will be classified as a lens in the test dataset.   We want to include these objects because we want to test the limits of the classifiers.  As we will see, if you restrict your objectives to detecting only lensed images with surface brightness above threshold, for example, the "best" algorithm might change and the TPR will change.  For this reason we plot the AUROC, TPR$_0$ and TPR$_{10}$ as a function of several variables for all the entries.  This is done by removing all the lenses that do not exceed the threshold and then recalculating the these quantities, the number of non-lenses remains the same.

\subsection{Performance of methods}
\label{sec:performance}

\begin{table*}
\centering
\begin{tabular}{llrrrl}
  \hline
  Name & type & AUROC & TPR$_0$ & TPR$_{10}$ & short description \\ 
  \hline
 CMU-DeepLens-ResNet-ground3 & Ground-Based & 0.98 & 0.09 & 0.45 & CNN \\ 
  CMU-DeepLens-Resnet-Voting & Ground-Based & 0.98 & 0.02 & 0.10 & CNN \\ 
  LASTRO EPFL & Ground-Based & 0.97 & 0.07 & 0.11 & CNN \\ 
  CAS Swinburne Melb & Ground-Based & 0.96 & 0.02 & 0.08 & CNN \\ 
  AstrOmatic & Ground-Based & 0.96 & 0.00 & 0.01 & CNN \\ 
  Manchester SVM & Ground-Based & 0.93 & 0.22 & 0.35 & SVM / Gabor \\ 
  Manchester-NA2 & Ground-Based & 0.89 & 0.00 & 0.01 & Human Inspection \\ 
   ALL-star & Ground-Based & 0.84 & 0.01 & 0.02 & edges/gradiants and Logistic Reg. \\ 
   CAST & Ground-Based & 0.83 & 0.00 & 0.00 & CNN / SVM \\ 
   YattaLensLite & Ground-Based & 0.82 & 0.00 & 0.00 & SExtractor \\ 
   LASTRO EPFL & Space-Based & 0.93 & 0.00 & 0.08 & CNN \\ 
  CMU-DeepLens-ResNet & Space-Based & 0.92 & 0.22 & 0.29 & CNN \\ 
   GAMOCLASS & Space-Based & 0.92 & 0.07 & 0.36 & CNN \\ 
  CMU-DeepLens-Resnet-Voting & Space-Based & 0.91 & 0.00 & 0.01 & CNN \\ 
  AstrOmatic & Space-Based & 0.91 & 0.00 & 0.01 & CNN \\ 
   CMU-DeepLens-ResNet-aug & Space-Based & 0.91 & 0.00 & 0.00 & CNN \\ 
   Kapteyn Resnet & Space-Based & 0.82 & 0.00 & 0.00 & CNN \\ 
   CAST & Space-Based & 0.81 & 0.07 & 0.12 & CNN \\ 
  Manchester1 & Space-Based & 0.81 & 0.01 & 0.17 & Human Inspection \\ 
   Manchester SVM & Space-Based & 0.81 & 0.03 & 0.08 & SVM / Gabor \\ 
   NeuralNet2 & Space-Based & 0.76 & 0.00 & 0.00 & CNN / wavelets \\ 
   YattaLensLite & Space-Based & 0.76 & 0.00 & 0.00 & Arcs / SExtractor \\ 
   All-now & Space-Based & 0.73 & 0.05 & 0.07 & edges/gradiants and Logistic Reg. \\ 
  GAHEC IRAP & Space-Based & 0.66 & 0.00 & 0.01 & arc finder \\ 
   \hline
\end{tabular}
\caption{The AUROC, TPR$_0$ and TPR$_{10}$ for the entries in order of AUROC.}
\label{table:AUROC}
\end{table*}

Table~\ref{table:AUROC} shows the AUROC, TPR$_0$ and TPR$_{10}$ for the entries in order of AUROC and dataset type.  It can be seen that CMU-DeepLens-ResNet-ground3 had the best AUROC for the ground-based set and LASTRO EPFL the best for the space-based set.  The order is different if TPR$_0$ is used to rank the entries as seen in table~\ref{table:TPR0}.  Here Manchester SVM and 
CMU-DeepLens-ResNet get the best scores.

\begin{table*}
\centering
\begin{tabular}{llrrrl}
  \hline
  Name & type & AUROC & TPR$_0$ & TPR$_{10}$ & short description \\ 
  \hline
 Manchester SVM & Ground-Based & 0.93 & 0.22 & 0.35 & SVM / Gabor \\ 
  CMU-DeepLens-ResNet-ground3 & Ground-Based & 0.98 & 0.09 & 0.45 & CNN \\ 
  LASTRO EPFL & Ground-Based & 0.97 & 0.07 & 0.11 & CNN \\ 
   CMU-DeepLens-Resnet-Voting & Ground-Based & 0.98 & 0.02 & 0.10 & CNN \\ 
   CAS Swinburne Melb & Ground-Based & 0.96 & 0.02 & 0.08 & CNN \\ 
   ALL-star & Ground-Based & 0.84 & 0.01 & 0.02 & edges/gradiants and Logistic Reg. \\ 
   Manchester-NA2 & Ground-Based & 0.89 & 0.00 & 0.01 & Human Inspection \\ 
   YattaLensLite & Ground-Based & 0.82 & 0.00 & 0.00 & SExtractor \\ 
   CAST & Ground-Based & 0.83 & 0.00 & 0.00 & CNN / SVM \\ 
   AstrOmatic & Ground-Based & 0.96 & 0.00 & 0.01 & CNN \\ 
   CMU-DeepLens-ResNet & Space-Based & 0.92 & 0.22 & 0.29 & CNN \\ 
   GAMOCLASS & Space-Based & 0.92 & 0.07 & 0.36 & CNN \\ 
   CAST & Space-Based & 0.81 & 0.07 & 0.12 & CNN \\ 
   All-now & Space-Based & 0.73 & 0.05 & 0.07 & edges/gradiants and Logistic Reg. \\ 
   Manchester SVM & Space-Based & 0.80 & 0.03 & 0.07 & SVM / Gabor \\ 
   Manchester1 & Space-Based & 0.81 & 0.01 & 0.17 & Human Inspection \\ 
   LASTRO EPFL & Space-Based & 0.93 & 0.00 & 0.08 & CNN \\ 
   GAHEC IRAP & Space-Based & 0.66 & 0.00 & 0.01 & arc finder \\ 
   AstrOmatic & Space-Based & 0.91 & 0.00 & 0.01 & CNN \\ 
   Kapteyn Resnet& Space-Based & 0.82 & 0.00 & 0.00 & CNN \\ 
   CMU-DeepLens-ResNet-aug & Space-Based & 0.91 & 0.00 & 0.00 & CNN \\ 
   CMU-DeepLens-Resnet-Voting & Space-Based & 0.91 & 0.00 & 0.01 & CNN \\ 
   NeuralNet2 & Space-Based & 0.76 & 0.00 & 0.00 & CNN / wavelets \\ 
   YattaLensLite & Space-Based & 0.76 & 0.00 & 0.00 & Arcs / SExtractor \\ 
   \hline
\end{tabular}
\caption{The AUROC, TPR$_0$ and TPR$_{10}$ for the entries in order of TPR$_0$.  }
\label{table:TPR0}
\end{table*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/roc_space.pdf}
 \caption{The ROC curves for the space-based entries.}
 \label{fig:roc_space}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/roc_ground.pdf}
 \caption{The ROC curves for the ground-based entries.  Notice that these are are generally better than in figure~\ref{fig:einstein_space} indicating that colour information is an important discriminant. }
 \label{fig:roc_ground}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/roc_kids.pdf}
 \caption{The ROC curves for the ground-based entries including only the cases with authentic images taken from the KiDS survey.  It can be seen that in all cases these are lower than in figure~\ref{fig:roc_ground}.}
 \label{fig:roc_kids}
\end{figure*}

Figures~\ref{fig:roc_space} and \ref{fig:roc_ground} show the ROC curves for all the entries.  In addition, figure~\ref{fig:roc_kids} shows the ROC curves for only the ground-based images where an actual KiDS image was used (see section~\ref{sec:sim-ground-based}).  It can be seen that the classifiers do uniformly less well on this subset.  This indicates that the simulated galaxy images are different from the real ones and that the classifiers are able to distinguish fake foreground galaxies from lenses more easily than from real galaxies.  This is perhaps not unexpected, but does show that the simulated lenses need to be improved before the raw numbers can be directly used to evaluate the performance of a classifier on real data.

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/einstein_space.pdf}
 \caption{The AUROC, TPR$_0$, TPR$_{10}$ and the fraction of lenses in the test sample after discarding the lenses with Einstein radii larger than the number indicated on the x-axis. }
 \label{fig:einstein_space}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/einstein_ground.pdf}
 \caption{Same as figure~\ref{fig:einstein_space}, but for space-based entries.}
 \label{fig:einstein_ground}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/flux_space.pdf}
 \caption{Same as figure~\ref{fig:einstein_space}, but here the x-axis is the flux within the pixels that are above 1 $\sigma$ in the lensed source only image.}
 \label{fig:flux_space}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/flux_ground.pdf}
 \caption{Same as figure~\ref{fig:flux_space}, but for space-based entries.}
 \label{fig:flux_ground}
\end{figure*}

\begin{figure*}
\includegraphics[width=2\columnwidth]{figures/npixel_space.pdf}
 \caption{Same as figure~\ref{fig:einstein_space}, but here the x-axis is the number of pixels that are above 1 $\sigma$ in the lensed source only image.  This is an indication of the lensed arcs' size.}
 \label{fig:npixel_space}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/npixel_ground.pdf}
 \caption{Same as figure~\ref{fig:npixel_space}, but for space-based entries.}
 \label{fig:npixel_ground}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/flux_contrast_space.pdf}
  \caption{Same as figure~\ref{fig:einstein_space}, but here the x-axis is the ratio of the flux coming from the lensed source to the total flux in the image.  (\red{ check this.  Band?})}
 \label{fig:flux_contrast_space}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/flux_contrast_ground.pdf}
 \caption{Same as figure~\ref{fig:flux_contrast_space}, but for space-based entries.}
 \label{fig:flux_contrast_ground}
\end{figure*}

Figures~\ref{fig:einstein_space} and \ref{fig:einstein_ground} shows the AUROC, TPR$_0$, TPR$_{10}$ and fraction of lenses as a function of a lower cutoff on the Einstein radius (area).  There is also a vertical dotted line that indicates where no more than 100 lenses in the test sample had Einstein radii larger than this mark.  Beyond this point one should be suspicious of small number statistics.  When deriving the distribution of Einstein radii from data these curves would need to be used to correct for detection bias.  It can be seen the CMU-DeepLens-ResNet, Manchester1, Manchester SVM and GAMOCLASS obtain significantly higher TPR$_0$ and  TPR$_{10}$ for larger Einstein radii.  Manchester1 is the human inspection method.  In some cases the TPR$_0$'s are above 50\% of the lenses that meet this criterion.  Remember that many of the so called lenses are very dim or there is no significant arc because of the source position is well outside the caustic.  If an additional require was placed on the definition of a lens, such as the brightness of the arc being above a threshold, the TPRs would go up.

Figures~\ref{fig:flux_space} and \ref{fig:flux_ground} are the same except that the flux in the lensed image is used as the threshold.  We count only the flux in pixels with flux over one $\sigma$ of the background.  In some cases one can see an abrupt rise in the TPRs at some flux threshold.  CMU-DeepLens-ResNet in particular reaches a TPR$_0$ above 75\% for the brightest $\sim$ 10\% of the lenses.

A lensed image can be bright without being visibly distorted as in the case of a unresolved images.  Figures~\ref{fig:npixel_space} and \ref{fig:npixel_ground} use the number of pixels in the lensed image(s) that are over  one $\sigma$ of the background.  In this case also some classifiers show an abrupt improvement when the image is required to be larger than some threshold.  Interestingly in some cases the TPRs go down with lensed image size after reaching a peak.  This could be because they are not differentiating the arcs from companion galaxies as well in this regime.  There were also cases where the arc intersects with the borders of the image that might cause them to be missed.

Figures~\ref{fig:flux_contrast_space} and \ref{fig:flux_contrast_ground} investigate how the flux contrast between the foreground objects and the lensed source affects the classifiers.  Interestingly some method's TPRs go up with this quantity and some go down.  \red{explanation?}

\red{surface brightness contrast}

\red{breakups Manchester between Neil and Amit}

\section{Conclusions \& discussion}
\label{sec:conclusion}

A large variety of lens finding methods were tested on simulated images that were developed separately.    We found that some methods could recover more than 50\% of the lenses above a lens image brightness or size threshold without a single false positive.  If the data closely resembled the simulations we would already have reasonably good methods whose efficiency and biases can be quantitatively characterized.

We have done a fairly good jobs of determining that lenses can be identified in a population of fairly "normal" galaxies.  It is the rare "abnormal" objects that pose the greatest challenge.  When real KiDS data was used in the simulations the classifiers were all less accurate and it was only human inspection that found the one jackpot lens in the data.  Things like ring galaxies, tidal tails in merging galaxies and irregular galaxies can be mistaken for lenses and were not well represented in the simulated data.  Accurately reproducing these objects will be an objective of future work.

It was surprising to some of the authors how well CNN and SVM methods did relative to human inspection.  These methods find differences in the classes of images that are not obvious to a human and can classify things as lenses with high confidence where a human would have doubt.  This ability comes with some danger of over fitting to the training set however.  The distinguishing characteristics might only be a property of simulated data and not of real data.  The confidence one will have in these methods is really the confidence one has in the realism of the simulations.

When initiating this project we had a concern that current methods would be too slow or require too much human intervention to handle large data sets.  Happily this seems not to be a problem with most of the automatic methods.  The CNN and SVM codes take some time to train, but once trained they are very fast in classifying objects.  Billions of objects can be easily handled.

Another lesson is that color information is very important.  Even with lower noise levels, higher resolution, a simpler PSF and no masking the lenses in the space-based set were harder to find than the lenses in the ground-based set.  Having multiple bands clearly makes a big difference.  Euclid will have several infrared bands with lower resolution than the visible images that were not included in the challenge.  Even rather low resolution information from another instrument or telescope when combined with higher resolution data in one band might significantly improve the detection rates.

\section*{Acknowledgements}

EB thanks Raphael Gavazzi for proposing him to participate to the challenge.
AS was supported by World Premier International Research Center Initiative (WPI Initiative), MEXT, Japan.
RBM's research was partly part of project GLENCO, funded under the European Seventh Framework Programme, Ideas, Grant Agreement n. 259349.
AT acknowledges receipt of an STFC postdoctoral research
assistantship.
\red{We thank the International Space Science Institute (ISSI) for hosting and funding our workshop.\footnote{http://www.issibern.ch/}}
CA acknowledges support from the Enrico Fermi
Institute at the University of Chicago, and the University of Chicago
Provost's Office. NL would like to thank the funding support from
NSFC, grant no.11503064, and Shanghai Natural Science Foundation,
grant no. 15ZR1446700. This work was also supported in part by the
Kavli Institute for Cosmological Physics at the University of Chicago
through grant NSF PHY-1125897 and an endowment from the Kavli
Foundation and its founder Fred Kavli.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

\bibliographystyle{mnras}
\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%

\appendix

\label{lastpage}
\end{document}